---
title: "ephys4"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ephys4}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Read, plot and analyse ephys data

The ephys4 package facilitates reading HEKA Patchmaster files into an R session, extract and plot single traces or complete series, and perform analyses on them.

We start by loading our library, *ephys4*. We also load a second library, *ephysdata*, which contains examplefiles that we will use here as demo data.

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
library(ephys4)
library(ephysdata)
library(patchwork)
```

```{r fake_setup, include=TRUE, eval=FALSE}

library(ephys4)
library(ephysdata)
library(patchwork)
```

Working with *ephys4* always starts with loading some data from one of the supported platforms. Let's load some HEKA patch clamp data:

```{r load1}
read_PATCHMASTER(examplefile("NaV"))
```

The result is a table showing an overview of all experiments in the file, one sweep per row. We can do all typical tidyverse operations on the table, like filter, mutate or summarise:

```{r}
read_PATCHMASTER(examplefile("NaV")) %>% 
  group_by(exp, ser, serlabel) %>% summarise(n_sweeps=n())
```

This gives us a nice overview what is in the file: we have 1 experiment, featuring different series with different names.

The last column, *ptrs*, contains a link to the actual raw data of the sweeps. This is needed to make a plot of the recorded data. This can be done using the `ggsweeps()` function. Let's select the first of the three series labeled "NaIV" and plot it:

```{r, out.width='100%', fig.width=12}
NaIV_1 <- read_PATCHMASTER(examplefile("NaV")) %>% 
  filter(exp==1, ser==2)  

NaIV_1 %>%  ggsweeps() + 
  scalebars( yunit="pA", xunit = "ms", yfactor=1000, xfactor=1000, expand = F, xlength = .001) +

NaIV_1 %>%  ggsweeps() + 
  scalebars(x=.7, y=.1,  yunit="pA", xunit = "ms", yfactor=1000, xfactor=1000, expand = F, xlength = .01) 
```

## Adding cursors

The most interesting feature is obviously the peak that occours near 0.01 seconds. How big is it ? To do this analysis, we add a cursor:

```{r , out.width='100%', fig.width=8}
peaks <- 
 NaIV_1 %>%
  add_cursor_point("peak" , start=0.009, end=0.013, min)  

print(peaks)
```

This gives us a modified version of the table, which now has a column "peak" showing us the values of the minimum in the specified region. If we pipe this table to the `ggsweeps()` function, the values are automatically shown:

```{r, out.width='100%', fig.width=12}
peaks  %>% 
  ggsweeps() + 
  theme_classic()
```

of course, we can also use the data in the table to do a simple ggplot directly:

```{r, out.width='100%', fig.width=12}
peaks %>% 
  ggplot(aes(swp.start, peak)) + 
  geom_line() + 
  geom_point() + 
  theme_classic()
```

## Add bar markers

```{r, out.width='100%', fig.width=12, fig.height=9}
peaks %>% filter(peak==min(peak)) %>%
  add_bar(name="prepulse",       start=0.00,   end=0.0098, line=1) %>% 
  add_bar(name="depolarisation", start=0.01,   end=0.03,   line=1) %>% 
  add_bar(name="recovery",       start=0.0302, end=0.04,   line=1) %>% 
  add_bar(name="late current",   start=0.015,  end=0.03,   line=2, fill="red", label.col="white") %>%
  add_bar(name="early peak",     start=0.010,  end=0.0148, line=2, border="black") %>%
  ggsweeps()+ 
  theme_classic()
```
